Throughput Prediction on 60 GHz Devices for Latency Sensitive Applications

### 2021-01-21 paper.tex
changes

In this work, {for the first time}, we study the throughput predictability in 60 GHz wireless LANs using machine learning. In the past, throughput estimation/prediction has been studied in the context of sub-6 GHz wireless networks~\cite{khan:infocom2016,song:secon2017,kajita:wmnc2015} and mobile networks~\cite{liu:globecom2015}. However there has been no previous work, to our best knowledge, which has studied throughput prediction at mmWave frequencies such as 60 GHz.

There are 2 main challenges we face to conduct this study. First, to reliably train and test any machine learning model, a significant amount of data is needed.~\cite{mao:sigcomm2017} used over 30 hours of network traces, ~\cite{yan:nsdi2020} collected data for over a million video streams, and~\cite{xu:conext2017} collected over 150 hours of video playback data through user trials. Since 60 GHz WLANs are not widely deployed (the first 802.11ad-enabled smartphones were only launched in 2019), we cannot collect data from real networks, unlike in ~\cite{mao:sigcomm2017,yan:nsdi2020}. Also, the only 2 phones that support 802.11ad are not VR Ready and so we could not perform real VR experiments with volunteers as well. Thus, we wanted a way to collect traces in a controlled environment efficiently and in an automated way while still considering the importance of performing realistic mobility patterns. To meet this conflicting requirements, we mounted the smartphone on a programmable 3-axis motion controller and slider typically used by professional photographers. Using this setup, we were able to collect more than 100 hours of traces while running different applications and performing various controlled and random mobility patterns. The second challenge is that, unlike previous works, which make predictions only on coarse-grained timescales (in the order of a few s) for ABR video streaming, we study prediction at timescales as low as 10 ms (targeting VR applications). Hence, we need ML models that strike the right balance between being accurate as well as being lightweight enough to run on mobile devices within such short timescales.

@@

\begin{itemize}
    \item 
\end{itemize}

>>>>

In this work, {for the first time}, we study the throughput predictability in 60 GHz wireless LANs using machine learning. In the past, throughput estimation/prediction has been studied in the context of sub-6 GHz wireless networks~\cite{khan:infocom2016,song:secon2017,kajita:wmnc2015} and mobile networks~\cite{liu:globecom2015}. However there has been no previous work, to our best knowledge, that has studied throughput prediction at mmWave frequencies such as 60 GHz.

There are 2 main challenges we face to conduct this study. First, to
reliably train and test any machine learning model, a significant
amount of data is needed.~\cite{mao:sigcomm2017} used over 30 hours of
network traces, ~\cite{yan:nsdi2020} collected data for over a million
video streams, and~\cite{xu:conext2017} collected over 150 hours of
video playback data through user trials. Since 60 GHz WLANs are not
widely deployed (the first 802.11ad-enabled smartphones were only
launched in 2019), we cannot collect data from real networks, unlike
in ~\cite{mao:sigcomm2017,yan:nsdi2020}. Also, the only 2 phones that
support 802.11ad are not VR Ready and so we could not perform real VR
experiments with volunteers. Thus, we wanted a way to collect
traces in a controlled environment efficiently and in an automated way
while still considering the importance of performing realistic
mobility patterns. To meet these conflicting requirements, we mounted
the smartphone on a programmable 3-axis motion controller and slider
typically used by professional photographers. Using this setup, we
were able to collect more than 100 hours of traces while running
different applications and performing various controlled and random
mobility patterns. The second challenge is that, unlike previous
works, which make predictions only on coarse-grained timescales (in
the order of a few seconds) for ABR video streaming, we study prediction at
timescales as low as 10 ms (targeting VR applications). Hence, we need
ML models that strike the right balance between being accurate as well
as being lightweight enough to run on mobile devices within such short
timescales.

@@

\begin{itemize}
    \item ???
\end{itemize}
